Reproducible Research  
=============================
*******************
Project1  
11/14/2015  
mdavis3419  

*****************
  
We are going to be examining what the walking pattern of an individual was over the course of October and November in 2012. We have data in 5 minute intervals recording the number of steps taken durring that period of time.

In order to work with this data we can move to the folder we want to work in.

```{r echo=FALSE}
setwd("C:\\Users\\Michael\\Documents\\GitHub\\Coursera\\Reproducable Research\\RepData_PeerAssessment1")
```
```
setwd("<your desired directory>")
```  

We can then retrieve the data by downloading, unziping, and loading the data from [my fork][1] of the [master GitHub repository][2] using the following function.

```{r}
getdata <- function(){
  temp <- tempfile()
  url <- "https://github.com/mdavis3419/RepData_PeerAssessment1/blob/master/activity.zip?raw=true"
  download.file(url, temp, mode="wb")
  unzip(temp, "activity.csv")
  read.table("activity.csv", sep=",", header=T)
  }
```

And we can assign the data to the variable CompleteActivityLog so that we can continue to work with the full dataset.

```{r}
  CompleteActivityLog <- getdata()
```

If we take a look at the first few rows of the data we can see that there are null values in the form of "NA", and that we have 3 fields "steps", "date", and "interval".

```{r}
  head(CompleteActivityLog)
```

There are two things we should do before we get started. First We are going to find out what the data types of each of the fields are.

```{r}
  sapply(CompleteActivityLog, class)
```

And next we are going to create a second dataset, which we will call nonNullActivityLog. Which only contains the values of the activity log that have non-null values in the steps column.

```{r}
  nonNullActivityLog <- CompleteActivityLog[!is.na(CompleteActivityLog),]
```

**************************

Now that we have all of our basic information, let's look at the first problem to solve.

### What is mean total number of steps taken per day?


## 
[1]: https://github.com/mdavis3419/RepData_PeerAssessment1 "my fork"
[2]: https://github.com/rdpeng/RepData_PeerAssessment1 "master GitHub repository"