Reproducible Research  
=============================
*******************
Project1  
11/14/2015  
mdavis3419  

*****************
  
We are going to be examining what the walking pattern of an individual was over the course of October and November in 2012. We have data in 5 minute intervals recording the number of steps taken durring that period of time.

In order to work with this data we can move to the folder we want to work in.

```{r echo=FALSE}
setwd("C:\\Users\\Michael\\Documents\\GitHub\\Coursera\\Reproducable Research\\RepData_PeerAssessment1")
```
```
setwd("<your desired directory>")
```  

We can then retrieve the data by downloading, unziping, and loading the data from [my fork][1] of the [master GitHub repository][2] using the following function.

```{r}
getdata <- function(){
  temp <- tempfile()
  url <- "https://github.com/mdavis3419/RepData_PeerAssessment1/blob/master/activity.zip?raw=true"
  download.file(url, temp, mode="wb")
  unzip(temp, "activity.csv")
  read.table("activity.csv", sep=",", header=T)
  }
```

And we can assign the data to the variable CompleteActivityLog so that we can continue to work with the full dataset.

```{r}
  CompleteActivityLog <- getdata()
```

If we take a look at the first few rows of the data we can see that there are null values in the form of "NA", and that we have 3 fields "steps", "date", and "interval".

```{r}
  head(CompleteActivityLog)
```

There are two things we should do before we get started. First We are going to find out what the data types of each of the fields are.

```{r}
  sapply(CompleteActivityLog, class)
```

And next we are going to create a second dataset, which we will call nonNullActivityLog. Which only contains the values of the activity log that have non-null values in the steps column.

```{r}
  nonNullActivityLog <- CompleteActivityLog[!is.na(CompleteActivityLog),]
```

**************************

Now that we have all of our basic information, let's look at the first problem to solve.

### What is mean total number of steps taken per day?

We are going to use the non-null dataset for this analysis so that we can get an understanding of the data we have records for.

We are going to approach this problem in a few steps.

First we are going to find the total number of steps per day.

```{r}
  stepsPerDay <- tapply(nonNullActivityLog$steps, nonNullActivityLog$date, sum)
  stepsPerDay
```

As you can see there are null dates in stepsPerDay. This is because the dataset was made dense to include days durring the activity months that we have no data for.
  
We can now create a histogram to show us the distribution of steps per day. While a look at the distribution at 5000 step buckets is good a better understanding of the shape of the data can be gained from a 2000 step bucket view.
  
```{r}
  buckets <- c(0,2000,4000,6000,8000,10000,12000,14000,16000,18000,20000,22000)
  histOfStepsPerDay <- hist(stepsPerDay, main="Histogram of Total Steps Per Day", xlab="2000 Step Buckets", ylab ="Frequency", breaks=buckets)
```

In addition to understanding the shape of this data, we should take a look at the mean and median of the data.

```{r}
  meanStepsPerDayExcludeNulls <- mean(stepsPerDay,na.rm = TRUE)
  meanStepsPerDayExcludeNulls
  medianStepsPerDayExcludeNulls <- median(stepsPerDay,na.rm = TRUE)
  medianStepsPerDayExcludeNulls
```

Now that we have a solid understanding of what typical daily walking habits of this individual were, let's take a look at how they walked throughout the day.

We are going to look at the mean number of steps they took durring the individual 5 minute intervals throughout the day.

```{r}
  meanStepsThroughoutDay <- tapply(nonNullActivityLog$steps, nonNullActivityLog$interval, mean)
  head(meanStepsThroughoutDay)
```

Now we are going to plot these averages to see how activity changes throughout the day.
```{r}
  plot(strtoi(names(meanStepsThroughoutDay))/100,meanStepsThroughoutDay, type="l", main = "Average Steps Taken Durring 5 Minute Intervals", xlab="Time of Day (24h)",ylab="Steps (#)", xaxt="n" )
  axis(1,at=seq(0,24, by=6),las=2)
```

We can see that some time between 8 and 9 AM there is a spike in the average number of steps taken. In order to find out
  
## 
[1]: https://github.com/mdavis3419/RepData_PeerAssessment1 "my fork"
[2]: https://github.com/rdpeng/RepData_PeerAssessment1 "master GitHub repository"